---
title: "prey_capture_Gasteracantha"
author: "Fabian C. Salgado-Roa"
date: "9/19/2022"
output:
  pdf_document:
    toc: yes
    toc_depth: '5'
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 5
    highlight: tango
    theme: paper
    self_contained: yes
    mode: selfcontained
editor_options:
  chunk_output_type: console
---

# Introduction

## Load the libraries and standard error function

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(glmmTMB)
library(car)
library(DHARMa)
```

Standard error function 

```{r}
se <- function(x) sd(x)/sqrt(length(x))
```

## Load the datasets

Total dataset

```{r, message=FALSE, warning=FALSE}
tmp_noNA<-read_csv("clean_data_and_scripts/data_gasteracantha_field.csv",col_names=T)
```

Because the measures are repeated for each individual, lets create a subset with a single observation to explore association between colour morph and the variables measured

```{r}
unique_ind <- tmp_noNA %>% distinct(code, .keep_all=T)
```

# Association between the predictors

Check the normality of the variables per colour category

```{r results='hide'}
#Web height
hist(unique_ind$Web_height)
qqPlot(unique_ind$Web_height,groups=factor(unique_ind$colour))
#opisthosoma width
hist(unique_ind$opisto_w)
qqPlot(unique_ind$opisto_w,groups=factor(unique_ind$colour))
#web Area
hist(log(unique_ind$web_area+10))
qqPlot(log(unique_ind$web_area+10),groups=factor(unique_ind$colour))
```

It seems that all the variables are following a normal distribution in every colour category

# Linear models

**Web height** 

Run the linear model

```{r}
mod_webHeight<-lm(Web_height~colour,data=unique_ind)
#plot(mod_webHeight)
```

Check the significance of the predictor

```{r}
summary(mod_webHeight)
anova(mod_webHeight)
```

**Opisthosoma width** 

Run the linear model

```{r}
mod_opistoW<-lm(unique_ind$opisto_w~colour,data=unique_ind)
#plot(mod_opistoW)
```

Check the significance of the predictor

```{r}
summary(mod_opistoW)
anova(mod_opistoW)
```

**Web area** 

Run the linear model

```{r}
mod_webArea<-lm(log(web_area+10)~colour,data=unique_ind)
#plot(mod_webArea)
```

Check the significance of the predictor

```{r}
summary(mod_webArea)
anova(mod_webArea)
```

*Because we found that the colour morphs differ in their opisthosoma width, we explored if this predictor is associated with prey capture*

**model with negative binomial distribution**

Run the model

```{r}
capture_opistoW<-glmmTMB(formula=total_preys ~ opisto_w + (1|code) + (1|check),
       family="nbinom2",REML=T,data=tmp_noNA)
```

Check the significance of the predictors

```{r}
Anova(capture_opistoW)
```


**model with zero inflated poisson**

Run the model

```{r}
capture_opistoW<-glmmTMB(formula=total_preys ~ opisto_w + (1|code) + (1|check),
       zi=~.,
       family="poisson",REML=T,data=tmp_noNA)
```

Check the significance of the predictors

```{r}
Anova(capture_opistoW)
Anova(capture_opistoW,component="zi")
```

**Web area and web height present collinearity**

Run the model

```{r}
mod_webArea_Heigth<-lm(log(web_area+10)~Web_height,data=unique_ind)
#plot(mod_webArea_Heigth)
```

Check the significance of the predictor

```{r}
summary(mod_webArea_Heigth)
Anova(mod_webArea_Heigth)
```

# Light differences between sides and colour morphs

**Light measures are bimodal, in consequences we decided to this variable as binary (high or low). This categorization can be explore **


```{r}
paleta<-c("#636363","#f0f0f0","#ffeda0")

ggplot(data=tmp_noNA, aes(x=log(luxes*100), group=colour, fill=colour)) +
    geom_density(adjust=1.5, alpha=.4) +
    scale_fill_manual(values=paleta)+
    scale_x_continuous(limits=c(0,12))+
    theme_classic()
```
**We explore if there were differences between sides in the exposure to a certain light condition**

Run the model

```{r}
light_side<-glmmTMB(formula=b_luxes ~ Side+ (1|code) + (1|check),family="binomial",data=tmp_noNA)
```

Check the significance of the predictor

```{r}
summary(light_side)
Anova(light_side)
```

We found that the ventral side is more often expose to conditions of high luminosity. Let's visualize the results

```{r}
#Predict the values based on the model
tmp_noNA$light_side<-predict(light_side,type="response")
#Colours
paleta_side<-c("#B84C7A","#BCA7D2")
#Plot
ggplot(data = tmp_noNA,aes(x=Side,y=light_side, fill=Side))+
scale_y_continuous(limits=c(0,1))+
     scale_fill_viridis_d( option = "D")+
       scale_fill_manual(values=paleta_side)+
     scale_colour_manual(values=paleta_side)+
     stat_summary(fun = mean,aes(color = Side,group=Side),fun.min = function(x) mean(x) - (2*se(x)),fun.max = function(x) mean(x)+(2*se(x)),geom = "pointrange", position = position_jitterdodge(jitter.width=0.15), size=1.5)+
     theme_classic()
```

**We also tested if the colour morphs differ in which side is expose to the different light environments**


Run the model with yellow and ventral as fixed levels

```{r}
tmp_noNA$colour<-relevel(as.factor(tmp_noNA$colour),"yellow")
tmp_noNA$Side<-relevel(as.factor(tmp_noNA$Side),"ventral")
colour_light1<-glmmTMB(formula=b_luxes ~ Side*colour + (1|code) + (1|check),family="binomial",data=tmp_noNA)
```

Check the significance of the predictor

```{r}
summary(colour_light1)
Anova(colour_light1,type=3)
```

Run the model with black and dorsal as fixed levels

```{r}
tmp_noNA$colour<-relevel(as.factor(tmp_noNA$colour),"black")
tmp_noNA$Side<-relevel(as.factor(tmp_noNA$Side),"dorsal")
colour_light2<-glmmTMB(formula=b_luxes ~ Side*colour + (1|code) + (1|check),
       family="binomial",data=tmp_noNA)
```

Check the significance of the predictors

```{r}
summary(colour_light2)
Anova(colour_light2,type=3)
```

Let's visualize the results

```{r}
#Predict the values based on the model
tmp_noNA$predicted_light<-predict(colour_light1,type="response")
#Colours
paleta2<-c("#636363","#ffeda0","#f0f0f0")
#Plot
ggplot(data = tmp_noNA,aes(x=Side,y=predicted_light, fill=Side))+
scale_y_continuous(limits=c(0,1))+
     scale_fill_viridis_d( option = "D")+
     #geom_point(data = pop_size,aes(x=colour,y=b_luxes, fill=side, size=value,shape=side), position = position_jitterdodge(jitter.width=0.05),alpha=1)+
       scale_fill_manual(values=paleta2)+
     scale_colour_manual(values=paleta2)+
     stat_summary(fun = mean,aes(color = colour,group=colour),fun.min = function(x) mean(x) - (2*se(x)),fun.max = function(x) mean(x)+(2*se(x)),geom = "pointrange", position = position_jitterdodge(jitter.width=0.15), size=1.5)+
     theme_classic()
```
#Side filtering and analyses (Make this look pretty)

we compared dorsal and ventral light measures of each individual, where the higher measure was considered as open space and the lower measure as vegetation. We did this for everyday of observation. The following code will create a new dataset.

```{r}
unique_ind2<-data.frame()
for(d in unique(tmp_noNA$day)){
       sub<-tmp_noNA %>% filter(day==d & Time==6.3)
       unique_ind2<-rbind(unique_ind2,sub)
}
side_data<-tibble(code=character(),side=character(),open_space=numeric(),day=numeric())
for(d in unique(unique_ind2$day)){
       filtered_day<-unique_ind2 %>% filter(day==d)
       tmp<-tibble(code=character(),side=character(),open_space=numeric(),day=numeric())
       for(ind in unique(filtered_day$code)) {
              dorsal<-filtered_day %>% filter(code==ind & Side=="dorsal")
              ventral<-filtered_day %>% filter(code==ind & Side=="ventral")
              if(nrow(ventral)==0 | nrow(dorsal)==0){next}
              else if(ventral$luxes>dorsal$luxes){
                     ventral<-tibble(code=ind,side="ventral",open_space=1,day=d)
                     dorsal<-tibble(code=ind,side="dorsal",open_space=0,day=d)
                     tmp<-rbind(tmp,ventral,dorsal)
              } else if(ventral$luxes<dorsal$luxes){
                     ventral<-tibble(code=ind,side="ventral",open_space=0,day=d)
                     dorsal<-tibble(code=ind,side="dorsal",open_space=1,day=d)
                     tmp<-rbind(tmp,ventral,dorsal)
              } 
       }
side_data<-rbind(side_data,tmp)
}

```

Let's evaluate if the new categorization is independent of how the spider position its body towards the different light conditions (High or Low luminosity)

```{r}
test_association<-chisq.test(factor(unique_ind2$t_luxes),factor(unique_ind2$site_exp))
test_association
```

Because we did not find association, let's run an independent model

```{r}
open_side<-glmmTMB(formula=open_space ~ side+ (1|code),family="binomial",data=side_data) #test which side is most
```

Check the model's fit

```{r , message=FALSE, warning=FALSE, results = 'hide'}
simulationOutput <- simulateResiduals(fittedModel = open_side, n = 1000)
testDispersion(simulationOutput = simulationOutput, alternative ="two.sided")
testUniformity(simulationOutput = simulationOutput, alternative ="two.sided")
```

Check the significance of the predictor

```{r}
summary(open_side)
Anova(open_side)
```

# Capture models

##Poisson model

Run the model with white and ventral as fixed levels

```{r}
tmp_noNA$colour<-relevel(as.factor(tmp_noNA$colour),"white")
tmp_noNA$Side<-relevel(as.factor(tmp_noNA$Side),"ventral")
total_model_poisson<-glmmTMB::glmmTMB(formula=total_preys ~ Side+colour+t_luxes+log(web_area) + (1|code) + (1|check),family="poisson",REML=F,data=tmp_noNA)
```

Test the fit of the model and zero inflation

```{r}
simulationOutput <- simulateResiduals(fittedModel = total_model_poisson, n = 1000)
testDispersion(simulationOutput = simulationOutput, alternative ="two.sided")
testUniformity(simulationOutput = simulationOutput, alternative ="two.sided")
testZeroInflation(simulationOutput = simulationOutput, alternative ="two.sided")
```

Because this model has zero inflation, dispersion and no uniformity; its is necessary to run a zero-inflation model to check is it fits better

```{r}
total_model_pzi<-glmmTMB::glmmTMB(formula=total_preys ~ colour+Side+t_luxes+log(web_area) + (1|code) + (1|check),zi=~.,family="poisson",REML=F,data=tmp_noNA)
```

Test the fit of the new model and zero inflation

```{r}
simulationOutput <- simulateResiduals(fittedModel = total_model_pzi, n = 1000)
testDispersion(simulationOutput = simulationOutput, alternative ="two.sided")
testUniformity(simulationOutput = simulationOutput, alternative ="two.sided")
testZeroInflation(simulationOutput = simulationOutput, alternative ="two.sided")
```

This model seems to be ok, let's now check the significance of the predictors

```{r}
summary(total_model_pzi)
Anova(total_model_pzi)
Anova(total_model_pzi,component="zi")
```

**We did not find any association between the predictor and the prey capture rate**

##Negative binomial model

Run the model with white and ventral as fixed levels

```{r}
total_model_nb<-glmmTMB::glmmTMB(formula=total_preys ~ colour+Side+t_luxes+log(web_area) + (1|code) + (1|check),family="nbinom2",REML=F,data=tmp_noNA)
```

Test the fit of the model and zero inflation

```{r}
simulationOutput <- simulateResiduals(fittedModel = total_model_nb, n = 1000)
testDispersion(simulationOutput = simulationOutput, alternative ="two.sided")
testUniformity(simulationOutput = simulationOutput, alternative ="two.sided")
testZeroInflation(simulationOutput = simulationOutput, alternative ="two.sided")
```

This model seems to be ok, let's now check the significance of the predictors

```{r}
summary(total_model_nb)
Anova(total_model_nb)
```

To create the plot fothe results, we have to predict the values under the model

```{r}
#Predict values to plot
tmp_noNA$response_nb<-predict(total_model_nb,type="response")
```

**With this, we can create two plots, one per side in order to make easy to visualize the results**

###Plot dorsal

Create subset

```{r}
plot_dorsal <- tmp_noNA %>% filter(Side=="dorsal")
plot_dorsal$colour<-relevel(as.factor(plot_dorsal$colour),"black")
```

Plot the results

```{r}
#Colours
paleta2<-c("#636363","#f0f0f0","#ffeda0")
#plot
ggplot(data = plot_dorsal,aes(x=t_luxes,y=log(response_nb+10), fill=colour))+
     scale_fill_viridis_d( option = "D")+
     geom_point(data = plot_dorsal,aes(x=t_luxes,y=log(total_preys+10), fill=colour),shape = 21,size=4, position = position_jitterdodge(jitter.width=0.15,jitter.height=0.01),alpha=1.0)+
       scale_fill_manual(values=paleta2)+
     scale_colour_manual(values=paleta2)+
     stat_summary(fun = mean,aes(color = colour,group=colour),fun.min = function(x) mean(x) - 2*se(x),fun.max = function(x) mean(x)+2*se(x),geom = "pointrange", position = position_jitterdodge(jitter.width=0.15), size=1.5)+
     theme_classic()

```

###Plot ventral

Create subset

```{r}
plot_ventral <- tmp_noNA %>% filter(Side=="ventral")
plot_ventral$colour<-relevel(as.factor(plot_ventral$colour),"black")
```

Plot the results

```{r}
#Colours
paleta2<-c("#636363","#f0f0f0","#ffeda0")
#plot
ggplot(data = plot_ventral,aes(x=t_luxes,y=log(response_nb+10), fill=colour))+
     scale_fill_viridis_d( option = "D")+
     geom_point(data = plot_ventral,aes(x=t_luxes,y=log(total_preys+10), fill=colour),shape = 21,size=4, position = position_jitterdodge(jitter.width=0.15,jitter.height=0.01),alpha=1.0)+
       scale_fill_manual(values=paleta2)+
     scale_colour_manual(values=paleta2)+
     stat_summary(fun = mean,aes(color = colour,group=colour),fun.min = function(x) mean(x) - 2*se(x),fun.max = function(x) mean(x)+2*se(x),geom = "pointrange", position = position_jitterdodge(jitter.width=0.15), size=1.5)+
     theme_classic()

```


